---
title: "Big Data and Global Surgery review"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(shiny)
library(plotly)
load("02_country_summary.rda")
load("02_country_data.rda")
load("02_allstudies.rda")
load("04_myworldmap.rda")


myplotdata = allstudies %>% 
	filter(author != "Global Burden of Disease Cancer Collaboration") %>% 
	mutate(n.groups = cut(n,
																	 breaks = c(300, 500, 1000, 2000, 5000, 10000, 428346)) %>% 
				 	fct_recode("300-500"        = "(300,500]",
				 						 "500-1,000"      = "(500,1e+03]",
				 						 "1000-2000"      = "(1e+03,2e+03]",
				 						 "2,000-5,000"    = "(2e+03,5e+03]",
				 						 "5,000-10,000"   = "(5e+03,1e+04]",
				 						 "10,000-428,346" = "(1e+04,4.28e+05]")) %>% 
	mutate(info = paste0(author, ". (", year, "): ", n %>% formatC(format = "f", big.mark = ",", digits = 0), " patients, ", surgical_speciality))
	
annual_summary = myplotdata %>%
	group_by(year) %>%
	summarise(n_year = sum(n))

speciality_summary = myplotdata %>%
	group_by(surgical_speciality) %>%
	summarise(n_speciality = sum(n))

```

### Number of patients per country (59 studies, excluding the 11 where individual numbers were not distinguisable or n<100)

```{r}

mycols = RColorBrewer::brewer.pal(9, 'PuRd')[4:9]

mapdata %>% 
		mutate(n_country.groups = cut(n_country,
																	 breaks = c(300, 500, 1000, 2000, 5000, 10000, 428346)) %>% 
				 	fct_recode("300-500"        = "(300,500]",
				 						 "500-1,000"      = "(500,1e+03]",
				 						 "1000-2000"      = "(1e+03,2e+03]",
				 						 "2,000-5,000"    = "(2e+03,5e+03]",
				 						 "5,000-10,000"   = "(5e+03,1e+04]",
				 						 "10,000-428,346" = "(1e+04,4.28e+05]")) %>% 
  ggplot(aes(x=long, y = lat, group = group, fill = n_country.groups)) +
  geom_polygon(colour = "black") +
  #theme_void() +
  coord_map("moll") +
  ggthemes::theme_map() +
  theme(legend.position = "bottom",
        #legend.justification = 'center',
        legend.text = element_text(size=10, face='bold'),
        legend.title = element_blank(),
        plot.caption = element_text(size=9, face='bold'),
        plot.title   = element_text(size=12))+
  guides(fill=guide_legend(ncol=2, reverse=TRUE)) +
	scale_fill_viridis_d(na.value = NA)

country_summary$country %>% unique() %>% length()

```



### Studies each year, coloured by N patients (69 studies - Excluding the GBD as it's not cohort/database/registry).

```{r}


	
p1 = myplotdata %>% 	
	mutate(dummy = 1) %>% 
	group_by(year) %>% 
	arrange(n) %>% 
	mutate(nth_study = cumsum(dummy)) %>% 
	ggplot(aes(x = year)) +
	#geom_col(colour = "blue", fill = NA) +
	geom_point(aes(y = nth_study, fill = n.groups, text = info), shape = 21, size = 4) +
	theme_classic() +
	geom_text(data = annual_summary, aes(y = -0.2, label = n_year %>% formatC(format = "f", big.mark = ",", digits = 0)), hjust = -0.3) +
	coord_flip() +
	scale_x_continuous(breaks = 2008:2018) +
	scale_y_continuous("Studies each year", limits = c(-0.2, 27)) +
	scale_fill_viridis_d("N patients") +
	theme(legend.position = "top") +
	xlab("")


ggplotly(p1, tooltip = c("text"))

```

### Cumulative number of patients 2008-2018

```{r}


add_2013 = data_frame(year = 2013, patients_peryear = 0)

myplotdata %>% 	
	group_by(year) %>% 
	summarise(patients_peryear = sum(n)) %>% 
	bind_rows(add_2013) %>% 
	ungroup() %>% 
	arrange(year) %>% 
	mutate(patients_cumulativeyear = cumsum(patients_peryear)) %>% 
	ggplot(aes(x = year, y = patients_cumulativeyear)) +
	geom_col(colour = "blue", fill = NA) +
	theme_classic() +
	coord_flip() +
	scale_x_continuous(breaks = 2008:2018)

```



### Studies per speciality, coloured by N patients (69 studies - Excluding the GBD as it's not cohort/database/registry).

```{r}

p2 = myplotdata %>% 	
	mutate(dummy = 1) %>% 
	group_by(surgical_speciality) %>% 
	arrange(n) %>% 
	mutate(nth_study = cumsum(dummy)) %>% 
	mutate(info = paste0(author, ". (", year, "): ", n %>% formatC(format = "f", big.mark = ",", digits = 0), " patients")) %>% 
	ggplot(aes(x = fct_infreq(surgical_speciality))) +
	#geom_col(colour = "blue", fill = NA) +
	geom_point(aes(y = nth_study, fill = n.groups, text = info), shape = 21, size = 3.5) +
	theme_classic() +
	geom_text(data = speciality_summary, aes(y = -0.2, label = n_speciality %>% formatC(format = "f", big.mark = ",", digits = 0)), hjust = -0.3) +
	coord_flip() +
	scale_y_continuous("N studies from 2008 to mid-2018", limits = c(-0.2, 27)) +
	scale_fill_viridis_d("N patients") +
	theme(legend.position = "top") +
	xlab("")


ggplotly(p2, tooltip = c("text"))

```


### Studies per outcome, coloured by N patients (69 studies - Excluding the GBD as it's not cohort/database/registry).

```{r}


	
p3 = myplotdata %>% 	
	mutate(dummy = 1) %>% 
	group_by(outcome_primary) %>% 
	arrange(n) %>% 
	mutate(nth_study = cumsum(dummy)) %>% 
	ggplot(aes(x = fct_infreq(outcome_primary))) +
	#geom_col(colour = "blue", fill = NA) +
	geom_point(aes(y = nth_study, fill = n.groups, text = info), shape = 21, size = 4) +
	theme_classic() +
	#geom_text(data = annual_summary, aes(y = -0.2, label = n_year %>% formatC(format = "f", big.mark = ",", digits = 0)), hjust = -0.3) +
	coord_flip() +
	#scale_x_continuous(breaks = 2007:2018) +
	scale_y_continuous("Studies each year", limits = c(-0.2, 27)) +
	scale_fill_viridis_d("N patients") +
	theme(legend.position = "top") +
	xlab("")


ggplotly(p3, tooltip = c("text"))

```



### Study size vs n studies per speciality


```{r}

compare_size_studies = myplotdata %>% 
	select(surgical_speciality, n) %>% 
	group_by(surgical_speciality) %>% 
	mutate(n_studies = n()) %>% 
	mutate(n_patients = sum(n)) %>% 
	mutate(n_patients.median = median(n)) %>% 
	distinct(surgical_speciality, n_studies, n_patients, n_patients.median)

# myplotdata$n %>% sum()
# compare_size_studies$n_studies %>% sum()
# compare_size_studies$n_patients %>% sum()

compare_size_studies %>% 
	gather(variable, value, -surgical_speciality) %>% 
	filter(variable != "n_patients.median") %>% 
	ggplot(aes(x  = surgical_speciality, y = value)) +
	geom_col() +
	facet_wrap(~variable, scales = "free") +
	coord_flip() +
	theme_bw()


```



### Study size vs n studies per speciality


```{r}
compare_size_studies %>% 
	ggplot(aes(x = n_studies, y = n_patients.median)) +
	geom_point() +
	geom_label(aes(label = surgical_speciality)) +
	theme_bw() +
	scale_y_continuous(expand = c(0.1, 0.1)) +
	scale_x_continuous(expand = c(0.1, 0.1))


```












