---
title: "Big Data and Global Surgery review"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(shiny)
library(plotly)
library(tidyverse)
library(knitr)
library(patchwork)
library(scales)
library(ggmap)
library(maptools)

#REFORMAT and TIDY the final included studies

allstudies = read_csv("final_included_papers_august2018.csv", skip = 19) %>% 
	mutate_if(is.character, str_trim) %>% 
	mutate(study_id = paste(author %>% tolower %>% str_replace_all(" ", "_"), year, sep = "-")) %>% 
	mutate(study_design = fct_recode(study_design,
																	 "Cohort (single centre)" = "Cohort (one centre)",
																	 "Cohort (country)"       = "Cohort (single country)") %>% 
				 	fct_relevel("Cohort (single centre)", "Cohort (country)", "Cohort (multiple countries)", "Database")) %>% 
	mutate(surgical_speciality = fct_recode(surgical_speciality,
																					"Breast surgery"        = "Breast",
																					"General surgery"       = "General",
																					"General surgery"       = "Urology",
																					"General surgery"       = "General Surgery",
																					"ENT"                   = "Head and neck",
																					"Multiple specialities" = "Multiple",
																					"Multiple specialities" = "Multiple specialities (elective surgery)")) %>% 
	mutate(surgical_speciality = fct_collapse(surgical_speciality,
																						"ENT/OMS"                    = c("ENT", "Maxillofacial surgery"),
																						"Obstetrics and gynaecology" = c("Obstetrics and gynaecology", "Gynaecology"))) %>% 
	mutate(outcome_primary = fct_recode(outcome_primary, "Patient characteristics" = "Patient demographics",
																			"Outcomes of surgery" = "Post-operative outcomes") %>% 
				 	fct_relevel("Patient characteristics", "Incidence", "Access", "Outcomes of surgery", "Patient cost", "Survival")) %>% 
	mutate(n_comment = fct_explicit_na(n_comment, "single")) %>% 
	mutate(n_comment.2groups = fct_collapse(n_comment,
																					"lmic_separated" = c("single", "multiple", "multiple_lmic"),
																					"lmic_hic_together" = c("multiple_worldwide", "multiple_hic"))) %>% 
	mutate(n.groups = cut(n,
												breaks = c(300, 500, 1000, 2000, 5000, 10000, 428346)) %>% 
				 	fct_recode("300-500"        = "(300,500]",
				 						 "500-1,000"      = "(500,1e+03]",
				 						 "1000-2000"      = "(1e+03,2e+03]",
				 						 "2,000-5,000"    = "(2e+03,5e+03]",
				 						 "5,000-10,000"   = "(5e+03,1e+04]",
				 						 "10,000-428,346" = "(1e+04,4.28e+05]")) %>% 
	mutate(info = paste0(author, ". (", year, "): ", n %>% formatC(format = "f", big.mark = ",", digits = 0), " patients, ", surgical_speciality))


# GET and COUNTRY information where possible
# reformat from wide to long

# str_extra() below
# extract number (in brackets after the country name), e.g. India (1000)
# str_remove() below
# now remove the number from the country name column, e.g. (1000) - digit in brackets



# available in the manuscripts in the Authorship+Collaborators section, using CSVs for convenience.

gs1_countries = read_csv("gs1_countries.csv") %>% 
	mutate(study_id = "globalsurg_collaborative-2016") %>% 
	rename(country = gs1)
gs2_countries = read_csv("gs2_countries.csv")  %>% 
	mutate(study_id = "globalsurg_collaborative-2018") %>% 
	rename(country = gs2)

gs_countries = bind_rows(gs1_countries, gs2_countries)

add_globalsurg_countries = allstudies %>% 
	filter(author == "GlobalSurg Collaborative") %>% 
	select(study_id, year, n_studytotal = n) %>% 
	full_join(gs_countries) %>% 
	mutate(n_country = NA)
rm(gs_countries, gs1_countries, gs2_countries)

# reshape other studies with multiple countries currently recorded in the same cell


lmic_countries = read_csv("wb_lmic_countries.csv") %>% 
	rename(country_wb = country)

countrylevel = allstudies %>% 
	filter(n_comment != "multiple_worldwide") %>% 
	select(study_id, year, n_studytotal = n, countries) %>% 
	separate(countries, into = paste0("country_", 1:30), sep = ",") %>% 
	gather(country_n, country, -study_id, -year, -n_studytotal) %>% 
	drop_na(country) %>% 
	select(-country_n) %>% 
	mutate(n_country = str_extract(country, "\\d+") %>% as.numeric())     %>%  
	mutate(country   = str_remove(country, "\\(\\d+\\)") %>% str_trim())  %>%    
	mutate(n_country = ifelse(is.na(n_country), n_studytotal, n_country)) %>% 
	mutate(n_country = ifelse(n_country == 999999999, NA, n_country)) %>% 
	bind_rows(add_globalsurg_countries) %>% 
	mutate(country_wb = fct_recode(country,
																 "Iran, Islamic Rep."   = "Iran",
																 "Congo, Dem. Rep."     = "Congo Dem. Rep.",
																 "Venezuela, RB"        = "Venezuela",
																 "Gambia, The"          = "Gambia",
																 "Egypt, Arab Rep."     = "Egypt",
																 "Congo, Rep."          = "Congo Rep.",
																 "Hong Kong SAR, China" = "Hong Kong, China (SAR)",
																 "West Bank and Gaza"   = "Palestine, State of",
																 "St. Kitts and Nevis"  = "Saint Kitts and Nevis"
	) %>% as.character()) %>% 
	left_join(lmic_countries)

rm(add_globalsurg_countries)

```

# Overview numbers

Final number of papers included in the systematic review: `r nrow(allstudies)` (see the full PRISMA diagram in the main paper.).

Some of these studies did not provide patient numbers split for their LMIC and HIC participants.

```{r total_n_studies}

allstudies %>% 
	add_count(n_comment.2groups, wt = n) %>% 
	select(-n) %>% 
	add_count(n_comment.2groups) %>% 
	distinct(n_comment.2groups, n, nn) %>% 
	select(description = n_comment.2groups, n_studies = n, n_patients = nn) %>% 
	kable()

```

Range of study sizes in the LMIC separated group:

```{r separate_lmic_studies}

lmic_separated = allstudies %>% 
	filter(n_comment.2groups == "lmic_separated")

lmic_separated$n %>% range()

```


Number of studies with over 3,000 LMIC patients:

```{r count_lmic_studies}

lmic_separated %>% 
	count(n > 3000) %>% 
	kable()
```

5 largest studies:


```{r top_5_studies}

lmic_separated %>% 
	arrange(-n) %>% 
	select(n, countries, study_design) %>% 
	slice(1:5) %>% 
	kable()

```


Studies where country-specific n patients was not separated/given:

```{r county_n_not_available}

countrylevel %>% 
	filter(is.na(n_country)) %>% 
	arrange(year) %>% 
	distinct(study_id) %>% 
	kable()

```



```{r sumamrise_countries}

countrysummary = countrylevel %>% 
	filter(lmic == "lmic") %>% 
	filter() %>% 
	group_by(country) %>% 
	summarise(n_studies = n_distinct(study_id),
						n_country = sum(n_country, na.rm = TRUE))
```


`r countrysummary$country %>% unique() %>% length()` LMIC countries have taken part in Surgical Big Data Studies.


# Yearly analysis (Figure X)


```{r yearly_plots}

# 0 studies from 2013 in our review so adding it in as a 0 for the cumulative plot
add_2013 = data_frame(year = 2013, patients_peryear = 0)

studies_peryear = lmic_separated %>% 
	count(year)

annual_summary = lmic_separated %>%
	count(year, wt = n) %>% 
	rename(patients_peryear = nn) %>% 
	bind_rows(add_2013) %>% 
	ungroup() %>% 
	arrange(year) %>% 
	mutate(patients_cumulativeyear = cumsum(patients_peryear))

p1 = lmic_separated %>% 
	mutate(dummy = 1) %>% 
	group_by(year) %>% 
	arrange(n) %>% 
	mutate(nth_study = cumsum(dummy)) %>% 
	ggplot(aes(x = year)) +
	#geom_col(colour = "blue", fill = NA) +
	geom_point(aes(y = nth_study, fill = n.groups, text = info), shape = 21, size = 3.5) +
	theme_classic() +
	coord_flip() +
	scale_x_continuous(breaks = 2008:2018) +
	scale_y_continuous("Studies each year", breaks = 1:max(studies_peryear$nn)) +
	scale_fill_viridis_d("N patients") +
	theme(legend.position = c(0.85, 0.2),
				axis.text = element_text(colour = "black")) +
	xlab("   ")


p2 = lmic_separated %>% 	
	group_by(year) %>% 
	summarise(patients_peryear = sum(n)) %>% 
	bind_rows(add_2013) %>% 
	ungroup() %>% 
	arrange(year) %>% 
	mutate(patients_cumulativeyear = cumsum(patients_peryear)) %>% 
	mutate(my_hjust = ifelse(year == 2018, 1.1, -0.1)) %>% 
	ggplot(aes(x = year, y = patients_cumulativeyear)) +
	geom_col(colour = "blue", fill = NA) +
	geom_text(aes(label = patients_cumulativeyear %>% 
									formatC(format = "f", big.mark = ",", digits = 0),
								hjust = my_hjust)) +
	theme_classic() +
	coord_flip() +
	scale_y_continuous(labels = comma, expand = c(0, 0)) +
	scale_x_continuous(breaks = 2008:2018) +
	theme(axis.text = element_text(colour = "black")) +
	xlab("") +
	ylab("Cumulative number of patients")

```


Column {.tabset  .tabset-fade}
-------------------------------------

### Interactive

```{r}

ggplotly(p1, tooltip = c("text"))
```   

### Print

```{r, fig.width=10}

p1+p2

```


# Maps (Figure 2)


```{r maps_setup}


countrysummary = countrysummary %>% 
	mutate(country = fct_recode(country,
															"Lao People's Democratic Republic" = "Lao PDR",
															"Burma"      = "Myanmar",
															"Viet Nam"   = "Vietnam",
															"Iran (Islamic Republic of)" = "Iran",
															"Democratic Republic of the Congo" = "Congo Dem. Rep.",
															"Congo" = "Congo Rep.",
															"Libyan Arab Jamahiriya" = "Libya",
															"Mauritania" = "Mauritius",
															"United Republic of Tanzania" = "Tanzania"
	) %>% as.character())


# Get map ------------
# https://groups.google.com/forum/#!msg/ggplot2/KAKhoE0GO4U/lsAXwQWNwhoJ
# Sundar Dorai-Raj
PolygonCoords <- function(polygon) {
	polygons <- polygon@Polygons
	coords.list <- lapply(seq_along(polygons), function(i) {
		# Extract the group, sequence, area, longitude, and latitude.
		coords <- polygons[[i]]@coords
		cbind(i, 1:nrow(coords), polygons[[i]]@area, coords)
	})
	coords.df <- as.data.frame(do.call(rbind, coords.list))
	names(coords.df) <- c("order", "seq", "area", "long", "lat")
	return(coords.df)
}

ConvertWorldSimple <- function(mapdata, min.area = 0) {
	# min.area is the minimum size of the polygons. Setting to some  # positive value will filter out tiny islands.
	coords.list <- lapply(mapdata@polygons, PolygonCoords)
	ncoords <- sapply(coords.list, nrow)
	coords.df <- do.call(rbind, coords.list)
	coords.df$country <- rep(mapdata@data$NAME, ncoords)
	country.group <- factor(paste(coords.df$country, coords.df$order))
	coords.df$group <- as.numeric(country.group)
	coords.df <- coords.df[coords.df$area >= min.area, ]
	return(coords.df)
}


#--------


data("wrld_simpl")
#remove small islands
world_map <- ConvertWorldSimple(wrld_simpl, min.area = 0.8) %>% 
	filter(country != "Antarctica")


# # check that names match
# countries_matching = world_map %>% 
# 	distinct(country) %>% 
# 	#rename(country = region) %>% 
# 	mutate(nameorigin = "mapdata")
# 
# mismatches = countrysummary %>% 
# 	left_join(countries_matching) %>% 
# 	filter(is.na(nameorigin))

mapdata = left_join(world_map, countrysummary) %>% 
	mutate(n_country.groups = cut(n_country,
																breaks = c(0, 100, 300, 1000, 3000, 15000, 40000, 473292))%>% 
				 	fct_recode("<100 or Unknown" = "(0,100]",
				 						 "100-300"         = "(100,300]",
				 						 "300-1,000"       = "(300,1e+03]",
				 						 "1,000-3,000"     = "(1e+03,3e+03]",
				 						 "1000-15,000"     = "(3e+03,1.5e+04]",
				 						 "15,000-40,000"   = "(1.5e+04,4e+04]" ,
				 						 "40,000-428,346"  = "(4e+04,4.73e+05]")) %>% 
	mutate(n_studies.groups = cut(n_studies, breaks = c(0, 1, 2, 3, 5, 10, 15)) %>% 
				 	fct_recode("1"     = "(0,1]",
				 						 "2"     = "(1,2]",
				 						 "3"     = "(2,3]",
				 						 "4-5"   = "(3,5]",
				 						 "6-10"  = "(5,10]",
				 						 "11-15" = "(10,15]"))


#mapdata$n_country.groups %>% levels()
#mapdata$n_studies.groups %>% levels()
```


```{r map_n_studies}

mycols = viridisLite::viridis(n = mapdata$n_country.groups %>% levels() %>% length())

m1 = mapdata %>% 
	ggplot(aes(x=long, y = lat, group = group, fill = n_studies.groups)) +
	geom_polygon(colour = "black") +
	#theme_void() +
	coord_map("moll") +
	ggthemes::theme_map() +
	theme(legend.position = "right",
				#legend.justification = 'center',
				legend.text = element_text(size=10, face='bold'),
				legend.title = element_blank(),
				plot.caption = element_text(size=9, face='bold'),
				plot.title   = element_text(size=12))+
	guides(fill=guide_legend(ncol=1, reverse=TRUE)) +
	scale_fill_viridis_d(na.value = NA) +
	ggtitle("(a) N studies per LMIC country")



```




```{r map_n_patients}

mycols = viridisLite::viridis(n = mapdata$n_country.groups %>% levels() %>% length())

m2 = mapdata %>% 
	ggplot(aes(x=long, y = lat, group = group, fill = n_country.groups)) +
	geom_polygon(colour = "black") +
	#theme_void() +
	coord_map("moll") +
	ggthemes::theme_map() +
	theme(legend.position = "right",
				#legend.justification = 'center',
				legend.text = element_text(size=10, face='bold'),
				legend.title = element_blank(),
				plot.caption = element_text(size=9, face='bold'),
				plot.title   = element_text(size=12))+
	guides(fill=guide_legend(ncol=1, reverse=TRUE)) +
	scale_fill_viridis_d(na.value = NA) +
	ggtitle("(b) N patients per LMIC country")




```



```{r, fig.width=20}

m1+m2

```



# Outcomes  (Figure 3)


```{r}

	
p3 = allstudies %>% 	
	mutate(dummy = 1) %>% 
	group_by(outcome_primary) %>% 
	arrange(n) %>% 
	mutate(nth_study = cumsum(dummy)) %>% 
	ggplot(aes(x = fct_rev(outcome_primary))) +
	geom_point(aes(y = nth_study, fill = n.groups, text = info), shape = 21, size = 4) +
	theme_classic() +
	coord_flip() +
	scale_y_continuous("Studies per outcome") +
	scale_fill_viridis_d("N patients") +
		theme(legend.position = c(0.85, 0.2),
				axis.text = element_text(colour = "black")) +
	guides(fill=guide_legend(ncol=4)) +
	xlab("")


```


Column {.tabset  .tabset-fade}
-------------------------------------

### Interactive

```{r}

ggplotly(p3, tooltip = c("text"))
```   

### Print

```{r, fig.width=10}

p3

```





